Predicting I/O-performance in HPC using Artificial Neural Networks

### 2016-07-03 Paper/paper.tex
weitere Bildanpassungen

The idea of the ema-model is to exploit periodic changes in the performance of the storage system.
Our analysis of measurements as well as the work of Crume et al. \cite{Crume:2013:FML:2538542.2538561} indicate that periodic phenomenons of access times in sequences of files accesses can be exploited for access time prediction.\\
Figure \ref{src/plot_From100001to100250_read_seq.png} shows a small series of measurements of sequentially read file accesses.
The peaks in access time are occurring periodically in this particular sequence. Such behavior appears only partially, often access times fluctuate without observable periodicity as in Figure.
An occurrence of a periodic pattern could be explained with a processing of the storage system in which a larger amount of data is loaded into the cache at the same time as soon as a cache-miss happens.
\fig{width=0.47\textwidth}{src/plot_From100001to100250_read_seq.png}{Small sequence of reading file accesses with periodic peaks in access time.}
\fig{width=0.47\textwidth}{src/plot_First250_read_seq.png}{Small sequence of reading file accesses with sporadic peaks in access time.}
With knowledge about these period of peaks a model could make better access time predictions.
Our second model is supposed to use its input information of previous data throughput for that.
The data throughput has peaks the same period as the access time.
Internally the model calculates the exponential moving average (EMA) of data throughput of measurement $i$ as:
\begin{equation}
EMA(i) = 0.5 \cdot t(i)+ 0.5 \cdot EMA(i-1)
\end{equation}
With $t(i)$ the access time of measurement $i$.
The influence of data throughput of a measurement is exponentially decreasing in the series of EMAs.
In Figure \ref{src/ema2.png} the EMA-function to a graph with periodic peaks can be seen.
\fig{width=.47\textwidth}{src/ema2.png}{Exponential moving average in red for a graph with perdiodic peaks.}\\
The ema-model can use $EMA(n-1)$ for the prediction of access time of measurement $n$.
If the model is able to find threshold values of the EMA-function after which another peaks follows, it can use this input information for better access time predictions on sequences with periodic peaks in access time. Depending on how relevant periodic behavior of access times is in our measurements, this additional input information could improve the model.\medskip

%The error-class-model is used as described in Section\,\ref{sec:error_classes} to examine the potential improvements for access time prediction due to knowledge about I/O-paths. TODO: wofür nochmal ?
% TODO: hier die Erkentnisse... ?? Welche Erkentnisse?

@@

% TODO: können wir die Datengröße noch dran schreiben? Einfach mit Gimp die Größe vertical drunter fügen anstelle Index ...
% TODO: Global fixen.
% Kann man in deren Style auch zwei BIlder nebeneinander zeichnen? Das wäre Cool. Ansonstne OK...

@@

\subsection{Prediction of file accesses}
After we analyzed the benchmarks and error classes we will study the results of our models.
To investigate the accuracy of the models and investigate overfitting, we split the available data into a training set and a test set. 
The test set consist of all measurements that weren't used for training.
The parameters for the ANN-models are varied to explore appropriate configurations; they are trained with a wide span of different parameter values for the number of hidden layers and number of neurons per layer.
They had 1000 randomly chosen measurements for each pair of access size and access type as training set. % TODO stimmt das 1000 pro Konfiguration?
In total that are 34\,000 measurements which should be a significant amount of data to avoid a situation in which we aren't able to train a model due to a lack of information about the system.

>>>>

The idea of the ema-model is to exploit periodic changes in the performance of the storage system.
Our analysis of measurements as well as the work of Crume et al. \cite{Crume:2013:FML:2538542.2538561} indicate that periodic phenomenons of access times in sequences of files accesses can be exploited for access time prediction.\\
Figure \ref{src/plot_From100001to100250_read_seq.png} shows a small series of measurements of sequentially read file accesses.
The peaks in access time are occurring periodically in this particular sequence. Such behavior appears only partially, often access times fluctuate without observable periodicity as in Figure.
An occurrence of a periodic pattern could be explained with a processing of the storage system in which a larger amount of data is loaded into the cache at the same time as soon as a cache-miss happens.
\fig{width=0.47\textwidth}{src/plot_From100001to100250_read_seq.png}{Small sequence of reading file accesses with periodic peaks in access time.}
\fig{width=0.47\textwidth}{src/plot_First250_read_seq.png}{Small sequence of reading file accesses with sporadic peaks in access time.}
With knowledge about these period of peaks a model could make better access time predictions.
Our second model is supposed to use its input information of previous data throughput for that.
The data throughput has peaks the same period as the access time.
Internally the model calculates the exponential moving average (EMA) of data throughput of measurement $i$ as:
\begin{equation}
EMA(i) = 0.5 \cdot t(i)+ 0.5 \cdot EMA(i-1)
\end{equation}
With $t(i)$ the access time of measurement $i$.
The influence of data throughput of a measurement is exponentially decreasing in the series of EMAs.
In Figure \ref{src/ema2.png} the EMA-function to a graph with periodic peaks can be seen.
\fig{width=.47\textwidth}{src/ema2.png}{Exponential moving average in red for a graph with perdiodic peaks.}\\
The ema-model can use $EMA(n-1)$ for the prediction of access time of measurement $n$.
If the model is able to find threshold values of the EMA-function after which another peaks follows, it can use this input information for better access time predictions on sequences with periodic peaks in access time. Depending on how relevant periodic behavior of access times is in our measurements, this additional input information could improve the model.

%The error-class-model is used as described in Section\,\ref{sec:error_classes} to examine the potential improvements for access time prediction due to knowledge about I/O-paths. TODO: wofür nochmal ?
% TODO: Soll hier nicht nochmal erwähnt werden, wofür das error-class-model genutzt wird?

@@

% TODO: können wir die Datengröße noch dran schreiben? Einfach mit Gimp die Größe vertical drunter fügen anstelle Index ...
% TODO: Global fixen.
% TODO Kann man in deren Style auch zwei BIlder nebeneinander zeichnen? Das wäre Cool. Ansonstne OK...

@@

\subsection{Prediction of file accesses}
After we analyzed the benchmarks and error classes we will study the results of our models.
To investigate the accuracy of the models and investigate overfitting, we split the available data into a training set and a test set. 
The test set consist of all measurements that weren't used for training.
The parameters for the ANN-models are varied to explore appropriate configurations; they are trained with a wide span of different parameter values for the number of hidden layers and number of neurons per layer.
They had 1000 randomly chosen measurements for each pair of access size and access type as training set. % TODO stimmt das 1000 pro Konfiguration? Ja, das stimmt :)
In total that are 34\,000 measurements which should be a significant amount of data to avoid a situation in which we aren't able to train a model due to a lack of information about the system.
